---
title: "sfcrime-timestamp"
author: "Antonio Marques Jr."
date: "9 September 2015"
output: html_document
---

#San Francisco Crimes Classification - Timestamp analysis

```{r, warning=FALSE, message=FALSE}
library(dplyr, quietly=TRUE)
library(ggplot2, quietly = TRUE)
sfcrime.timestamp <- read.csv("~/Documents/kaggle/sfcrime/sfcrime-dates/sfcrime-timestamp.csv")
sfcrime.timestamp <- sfcrime.timestamp %>% select(-seconds)
sfcrime.timestamp <- sfcrime.timestamp %>% filter(year != 2015)
summary(sfcrime.timestamp)
```

###YEAR

```{r}
summary(sfcrime.timestamp$year)
unique(sfcrime.timestamp$year)
```


```{r}
ggplot(sfcrime.timestamp, aes(x=factor(year))) + geom_histogram() + theme_bw()
```


###2003
```{r}
year_2003 <- sfcrime.timestamp %>% filter(year == 2003)
nrow(year_2003)
ggplot(year_2003, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2004
```{r}
year_2004 <- sfcrime.timestamp %>% filter(year == 2004)
nrow(year_2004)
ggplot(year_2004, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2005
```{r}
year_2005 <- sfcrime.timestamp %>% filter(year == 2005)
nrow(year_2005)
ggplot(year_2005, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2006
```{r}
year_2006 <- sfcrime.timestamp %>% filter(year == 2006)
nrow(year_2006)
ggplot(year_2006, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2007
```{r}
year_2007 <- sfcrime.timestamp %>% filter(year == 2007)
nrow(year_2007)
ggplot(year_2007, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2008
```{r}
year_2008 <- sfcrime.timestamp %>% filter(year == 2008)
nrow(year_2008)
ggplot(year_2008, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2009
```{r}
year_2009 <- sfcrime.timestamp %>% filter(year == 2009)
nrow(year_2009)
ggplot(year_2009, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2010
```{r}
year_2010 <- sfcrime.timestamp %>% filter(year == 2010)
nrow(year_2010)
ggplot(year_2010, aes(x=factor(month))) + geom_histogram() + theme_bw()
```


###2011
```{r}
year_2011 <- sfcrime.timestamp %>% filter(year == 2011)
nrow(year_2011)
ggplot(year_2011, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2012
```{r}
year_2012 <- sfcrime.timestamp %>% filter(year == 2012)
nrow(year_2012)
ggplot(year_2012, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2013
```{r}
year_2013 <- sfcrime.timestamp %>% filter(year == 2013)
nrow(year_2013)
ggplot(year_2013, aes(x=factor(month))) + geom_histogram() + theme_bw()
```

###2014
```{r}
year_2014 <- sfcrime.timestamp %>% filter(year == 2014)
nrow(year_2014)
ggplot(year_2014, aes(x=factor(month))) + geom_histogram() + theme_bw()
```


###Seasons: Spring, Summer, Fall, Winter

```{r}
sfcrime.timestamp$season <- ifelse((sfcrime.timestamp$month > 2) & (sfcrime.timestamp$month < 6),
                                   "spring",
                            ifelse((sfcrime.timestamp$month > 5) & (sfcrime.timestamp$month < 9),
                                   "summer",
                            ifelse((sfcrime.timestamp$month > 8) & (sfcrime.timestamp$month < 12),
                                   "fall", "winter")))
```


###Number of Crimes by Season

```{r}
ggplot(sfcrime.timestamp, aes(x=season)) + geom_histogram() + theme_bw()
```

```{r}
season_crimes_count <- sfcrime.timestamp %>% group_by(season) %>% summarise(count=n())
```


###Crimes in Spring
```{r}
crimes_spring <- sfcrime.timestamp %>% filter(season == "spring")
ggplot(crimes_spring, aes(x=factor(month))) + geom_histogram() + theme_bw()
crimes_spring_count <- crimes_spring %>% group_by(month) %>% summarise(count = n())
```

###Crimes in Fall

```{r}
crimes_fall <- sfcrime.timestamp %>% filter(season == "fall")
ggplot(crimes_fall, aes(x=factor(month))) + geom_histogram() + theme_bw()
crimes_fall_count <- crimes_fall %>% group_by(month) %>% summarise(count = n())
```

##Safer Seasons

###Crimes in Summer

```{r}
crimes_summer <- sfcrime.timestamp %>% filter(season == "summer")
ggplot(crimes_summer, aes(x=factor(month))) + geom_histogram() + theme_bw()
crimes_summer_count <- crimes_summer %>% group_by(month) %>% summarise(count = n())

```

###Crimes in Winter 
```{r}
crimes_winter <- sfcrime.timestamp %>% filter(season == "winter")
ggplot(crimes_winter, aes(x=factor(month))) + geom_histogram() + theme_bw()
crimes_winter_count <- crimes_winter %>% group_by(month) %>% summarise(count = n())
```


###Crimes by month - Count

```{r}
crimes_by_month_count <- rbind(crimes_winter_count, crimes_spring_count, crimes_summer_count, crimes_fall_count)

crimes_by_month_count <- crimes_by_month_count %>% arrange(-count)

```



###is_winter variable
```{r}
sfcrime.timestamp$is_winter <- ifelse(sfcrime.timestamp$season == "winter", TRUE, FALSE)
```

###fall_ending
```{r}
sfcrime.timestamp$fall_ending <- ifelse(sfcrime.timestamp$month == 10, TRUE, FALSE)

```

###spring_ending
```{r}
sfcrime.timestamp$spring_ending <- ifelse(sfcrime.timestamp$month == 5, TRUE, FALSE)
```


#Days in the months

###January
```{r}
january <- sfcrime.timestamp %>% filter(month == 1)
ggplot(january, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###Febryary
```{r}
february <- sfcrime.timestamp %>% filter(month == 2)
ggplot(february, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###March
```{r}
march <- sfcrime.timestamp %>% filter(month == 3)
ggplot(march, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###April
```{r}
april <- sfcrime.timestamp %>% filter(month == 4)
ggplot(april, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###May
```{r}
may <- sfcrime.timestamp %>% filter(month == 5)
ggplot(may, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###June
```{r}
june <- sfcrime.timestamp %>% filter(month == 6)
ggplot(june, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###July
```{r}
july <- sfcrime.timestamp %>% filter(month == 7)
ggplot(july, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###August
```{r}
august <- sfcrime.timestamp %>% filter(month == 8)
ggplot(august, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###September
```{r}
september <- sfcrime.timestamp %>% filter(month == 9)
ggplot(september, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###October
```{r}
october <- sfcrime.timestamp %>% filter(month == 10)
ggplot(october, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###November
```{r}
november <- sfcrime.timestamp %>% filter(month == 11)
ggplot(november, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###December
```{r}
december <- sfcrime.timestamp %>% filter(month == 12)
ggplot(december, aes(x=factor(day))) + geom_histogram() + theme_bw()
```

###Days in month
```{r}
days <- sfcrime.timestamp %>% group_by(day) %>% summarise(count=n())

ggplot(days, aes(x=factor(day), y=count, fill=count)) + geom_bar(stat="identity") + theme_bw()

```

```{r}

sfcrime.timestamp$is_thirty_first <- ifelse(sfcrime.timestamp$day == 31, TRUE, FALSE)
sfcrime.timestamp$is_first_of_month <- ifelse(sfcrime.timestamp$day == 1, TRUE, FALSE)
```


###Hours

```{r}
hours <- sfcrime.timestamp %>% group_by(hour) %>% summarise(count=n()) %>% arrange(-count)
ggplot(hours, aes(x=factor(hour), y=count, fill=count)) + geom_histogram(stat="identity") + theme_bw()
mean(hours$count)

sfcrime.timestamp$safe_hour <- ifelse( sfcrime.timestamp$hour >= 1 & sfcrime.timestamp$hour <= 7, TRUE, FALSE)

```

###MINUTES
```{r}
minutes <- sfcrime.timestamp %>% group_by(minutes) %>% summarise(count=n()) %>% arrange(-count)
ggplot(minutes, aes(x=factor(minutes), y=count, fill=count)) + geom_histogram(stat="identity") + theme_bw() 
sfcrime.timestamp$minutes_of_crime <- ifelse(sfcrime.timestamp$minutes == 0 | sfcrime.timestamp$minutes == 30, TRUE, FALSE)
```


```{r}
str(sfcrime.timestamp)

write.table(sfcrime.timestamp, "sfcrime-timestamp-2.csv", sep=",", quote = FALSE)


```